---
title: "Getting Started with piiR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with piiR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is the Predictive Information Index (PII)?

The Predictive Information Index (PII) quantifies how much outcome-relevant information is retained when reducing a set of predictors (e.g., items) to a composite score.

PII is defined as:

```math
\text{PII} = 1 - \frac{\text{Var}(\hat{Y}_{\text{Full}} - \hat{Y}_{\text{Score}})}{\text{Var}(\hat{Y}_{\text{Full}})}
```

Where:
- \( \hat{Y}_{\text{Full}} \): predictions from a full model (e.g., all items or predictors)
- \( \hat{Y}_{\text{Score}} \): predictions from a reduced score (e.g., mean or sum)

A PII of 1 means no predictive information was lost. A PII near 0 means the score loses most predictive information.

## Example: Using `pii()`

```{r}
library(piiR)

# Simulate two prediction vectors
set.seed(123)
full_model_preds <- rnorm(100)
score_based_preds <- full_model_preds + rnorm(100, sd = 0.5)

# Compute PII
pii(full_model_preds, score_based_preds)
```

